### 2015-10-30 pics/SampledTree1.tex
sampled ancestor tree pics

\usetikzlibrary{shapes,snakes}

\begin{tikzpicture}[thick, scale=0.6, color=gray]

\begin{scope}

%terminal nodes
\node at (-3.5, 2.5)(s2){};
\node at (0, 1)(s3){};
\node at (-2.5, 1)(s4){};
\node at (4, 3.1)(s5){};
\node at (1.4, 3.5)(s7){};
\node at (-1, 1)(s8){};
\node at (3.2, 1)(s9){};
\node at (2, 1)(s10){};

%branches
\draw (0, 7.5) -- (0,7) -- (-2, 7) --(-2, 5.5) --(-3, 5.5) -- (-3, 4.9) -- (-3.5, 4.9) -- (s2);
\draw (-3, 4.9) -- (-2.5, 4.9) -- (s4); 
\draw (-2, 5.5) -- (-1, 5.5) -- (s8); 
\draw (0,7) -- (2,7) -- (2,5.8) -- (0.7, 5.8) -- (0.7, 4.5) -- (0, 4.5) -- (s3); 
\draw (0.7, 4.5) -- (1.4, 4.5) -- (s7); 
\draw (2, 5.8) -- (3.3, 5.8) -- (3.3, 4) -- (2.6, 4) -- (2.6, 2.1) -- (2, 2.1) -- (s10); 
\draw (2.6, 2.1) -- (3.2, 2.1) -- (s9);
\draw (3.3, 4) -- (4, 4) -- (s5);

\node at (0, -1) {Full tree};

\pause

%sampled nodes
\node[fill,circle, inner sep=2pt] at (-2.5, 1)(sA){};
\node[anchor=north] (Alet) at (sA) {\textcolor{black}{\footnotesize A}};
\node[fill,circle, inner sep=2pt] at (0, 1)(sB){};
\node[anchor=north] (Alet) at (sB) {\textcolor{black}{\footnotesize B}};
\node[fill,circle, inner sep=2pt] at (2, 1)(sC){};
\node[anchor=north] (Alet) at (sC) {\textcolor{black}{\footnotesize C}};
\node[fill,circle, inner sep=2pt] at (3.2, 1)(sD){};
\node[anchor=north] (Alet) at (sD) {\textcolor{black}{\footnotesize D}};

\pause

\node[fill,circle, inner sep=2pt] at (3.3, 5)(sE){};
\node[anchor=west] (Alet) at (sE) {\textcolor{black}{\footnotesize E}};
\node[fill,circle, inner sep=2pt] at (-1, 2.9)(sF){};
\node[anchor=west] (Alet) at (sF) {\textcolor{black}{\footnotesize F}};

\pause

\color{newblue1}

%branches
\draw (0, 7.5) -- (0,7) -- (-2, 7) --(-2, 5.5) --(-3, 5.5) -- (-3, 4.9);
\draw (-3, 4.9) -- (-2.5, 4.9) -- (s4); 
\draw (-2, 5.5) -- (-1, 5.5) -- (sF); 
\draw (0,7) -- (2,7) -- (2,5.8) -- (0.7, 5.8) -- (0.7, 4.5) -- (0, 4.5) -- (s3); 
\draw (2, 5.8) -- (3.3, 5.8) -- (3.3, 4) -- (2.6, 4) -- (2.6, 2.1) -- (2, 2.1) -- (s10); 
\draw (2.6, 2.1) -- (3.2, 2.1) -- (s9);

%sampled nodes
\node[fill,circle, inner sep=2pt] at (-2.5, 1)(sA){};
\node[anchor=north] (Alet) at (sA) {\textcolor{black}{\footnotesize A}};
\node[fill,circle, inner sep=2pt] at (0, 1)(sB){};
\node[anchor=north] (Alet) at (sB) {\textcolor{black}{\footnotesize B}};
\node[fill,circle, inner sep=2pt] at (2, 1)(sC){};
\node[anchor=north] (Alet) at (sC) {\textcolor{black}{\footnotesize C}};
\node[fill,circle, inner sep=2pt] at (3.2, 1)(sD){};
\node[anchor=north] (Alet) at (sD) {\textcolor{black}{\footnotesize D}};

\node[fill,circle, inner sep=2pt] at (3.3, 5)(sE){};
\node[anchor=west] (Alet) at (sE) {\textcolor{black}{\footnotesize E}};
\node[fill,circle, inner sep=2pt] at (-1, 2.9)(sF){};
\node[anchor=west] (Alet) at (sF) {\textcolor{black}{\footnotesize F}};

\end{scope}

\begin{scope}[xshift=8.5cm]

\color{white}

%terminal nodes
\node[fill,circle, inner sep=2pt] at (-3, 1)(sA){};
\node[anchor=north] (Alet) at (sA) {\textcolor{white}{\footnotesize A}};
\node[fill,circle, inner sep=2pt] at (0.7, 1)(sB){};
\node[anchor=north] (Alet) at (sB) {\textcolor{white}{\footnotesize B}};
\node[fill,circle, inner sep=2pt] at (2.7, 1)(sC){};
\node[anchor=north] (Alet) at (sC) {\textcolor{white}{\footnotesize C}};
\node[fill,circle, inner sep=2pt] at (3.9, 1)(sD){};
\node[anchor=north] (Alet) at (sD) {\textcolor{white}{\footnotesize D}};
\node[fill,circle, inner sep=2pt] at (-1, 2.9)(sF){};
\node[anchor=west] (Alet) at (sF) {\textcolor{white}{\footnotesize F}};
\node[fill,circle, inner sep=2pt] at (3.3, 5)(sE){};
\node[anchor=west] (Alet) at (sE) {\textcolor{white}{\footnotesize E}};

%branches
\draw (0, 7.5) -- (0,7) -- (-2, 7) --(-2, 5.5) --(-3, 5.5)  -- (sA);
\draw (-2, 5.5) -- (-1, 5.5) -- (sF); 
\draw (0,7) -- (2,7) -- (2,5.8) -- (0.7, 5.8) -- (0.7, 4.5) -- (sB); 
\draw (2, 5.8) -- (3.3, 5.8) -- (3.3, 4) -- (3.3, 2.1) -- (2.7, 2.1) -- (sC); 
\draw (3.3, 2.1) -- (3.9, 2.1) -- (sD);

\end{scope}

\end{tikzpicture}